### 2024-04-19 docs/SOSP 2024/caching.tex
Shrink some figures
Fix some typos

The design of the \Treebeard{} compiler allows the implementation of different strategies for 
the in-memory representation of the model. The compiler currently has implementations for 
the three representations shown in Figure \ref{Fig:Representations}. The array and sparse 
representations are the ones described in the \TreebeardOLD{} paper\cite{Treebeard}. The reorg 
representation is the representation used by the RAPIDs library\cite{FIL}. 
The \textbf{array representation} is the simplest representation where the trees are stored in an array
in level order. The \textbf{sparse representation} stores the trees in a sparse format where 
memory is allocated only for nodes present in the tree and nodes contain pointers to their children. The \textbf{reorg representation}
interleaves the array representation of each tree in the model: all root nodes are stored first, then 
the left children of all the roots and so on. This representation was 
designed to improve memory coalescing when tree nodes are being loaded. 
\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/representations.png}
  \caption{The three representations supported by \Treebeard{}.}
  \label{Fig:Representations}
\end{figure}

One of the major changes we make to the original design of \TreebeardOLD{} \cite{Treebeard} is to
separate the implementation of representations from the rest of the compiler. This allows
us to implement representations as plugins to the compiler. We define an interface 
that representations implement. The code generator is implemented using
this interface thus hiding details of the actual representation from the core 
compiler. Curcially, the interface abstracts how and what buffers 
are allocated, how to move from a node to its child, how trees 
are cached, reading the value of leaves and now threshold and 
feature indices are read from the allocated buffers.
% The interface abstracts the following details. 
% \begin{itemize}
%   \item \textbf{Buffer allocation:} The representation object exposes methods that generate 
%   buffer allocations in the IR. 
%   \item \textbf{Moving to child nodes:} Given the value of the predicate at a node (or tile),
%   generate code to move to the appropriate child node of the current node.
%   \item \textbf{Leaf representation:} The interface abstracts determining whether the current
%   node is a leaf (which is needed to terminate walks) and how to get the value of the leaf
%   (which is needed to get a tree's prediction).
%   \item \textbf{Caching trees:} Since reading the trees into shared memory on GPU (or prefetching)
%   them on CPU require knowledge of the buffer layout, the task of generating caching code for trees
%   is delegated to the representation object. 
%   \item \textbf{Loading thresholds and feature indices:} The representation object provides MLIR 
%   rewrite patterns to lower operations that load thresholds and feature indices to LLVM IR. This
%   is necessary because the details of how to load the thresholds and feature indices from the model 
%   buffers is representation-specific.
% \end{itemize}

@@

%\subsection{Lowering of Caching Operations}
When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
the two caching operations is lowered differently for the CPU and the GPU. On CPU, the cahce operations are lowered to preteches while
on the GPU they are lowered to reads into shared memory.

>>>>

The design of the \Treebeard{} compiler allows the implementation of different strategies for 
the in-memory representation of the model. The compiler currently has implementations for 
the three representations shown in Figure \ref{Fig:Representations}. The array and sparse 
representations are the ones described in the \TreebeardOLD{} paper\cite{Treebeard}. The reorg 
representation is the representation used by the RAPIDs library\cite{FIL}. 
The \textbf{array representation} is the simplest representation where the trees are stored in an array
in level order. The \textbf{sparse representation} stores the trees in a sparse format where 
memory is allocated only for nodes present in the tree and nodes contain pointers to their children. The \textbf{reorg representation}
interleaves the array representation of each tree in the model: all root nodes are stored first, then 
the left children of all the roots and so on. This representation was 
designed to improve memory coalescing when tree nodes are being loaded. 
\begin{figure}[htb]
  \centering
  \includegraphics[width=0.7\linewidth]{figures/representations.png}
  \caption{The three representations supported by \Treebeard{}.}
  \label{Fig:Representations}
\end{figure}

One of the major changes we make to the original design of \TreebeardOLD{} \cite{Treebeard} is to
separate the implementation of representations from the rest of the compiler. This allows
us to implement representations as plugins to the compiler. We define an interface 
that representations implement. The code generator is implemented using
this interface thus hiding details of the actual representation from the core 
compiler. Crucially, the interface abstracts how and what buffers 
are allocated, how to move from a node to its child, how trees 
are cached, reading the value of leaves and now threshold and 
feature indices are read from the allocated buffers.
% The interface abstracts the following details. 
% \begin{itemize}
%   \item \textbf{Buffer allocation:} The representation object exposes methods that generate 
%   buffer allocations in the IR. 
%   \item \textbf{Moving to child nodes:} Given the value of the predicate at a node (or tile),
%   generate code to move to the appropriate child node of the current node.
%   \item \textbf{Leaf representation:} The interface abstracts determining whether the current
%   node is a leaf (which is needed to terminate walks) and how to get the value of the leaf
%   (which is needed to get a tree's prediction).
%   \item \textbf{Caching trees:} Since reading the trees into shared memory on GPU (or prefetching)
%   them on CPU require knowledge of the buffer layout, the task of generating caching code for trees
%   is delegated to the representation object. 
%   \item \textbf{Loading thresholds and feature indices:} The representation object provides MLIR 
%   rewrite patterns to lower operations that load thresholds and feature indices to LLVM IR. This
%   is necessary because the details of how to load the thresholds and feature indices from the model 
%   buffers is representation-specific.
% \end{itemize}

@@

%\subsection{Lowering of Caching Operations}
When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
the two caching operations is lowered differently for the CPU and the GPU. On CPU, the cache operations are lowered to prefetches while
on the GPU they are lowered to reads into shared memory.
